### 2021-06-29 case-study/Munging_CCES.Rmd
writeup on working with large surveys

---
title: "Processing large surveys with package MRP-KIT"
author: "Mitzi Morris"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE)
```

<style type="text/css">
.table { width: 40%; }
div.quote-container blockquote {  font-size: small; }
</style>

The goal of this case study is to show how the MRP-KIT package
can facilitate analysis of a large-scale survey sample.
By large-scale, we mean surveys which have both a rich set
of demographics and a large number of responses.
The richer the set of demographics, the larger the number
of possible sub-populations of interest; no matter how large
or representative the survey, the combinatorics of the demographic
categories are such that it is unlikely that the sample
is representative for all subpopulations of interest.
In order to correct this imbalance,
we post-stratify to a truly representative
sample drawn from the general population.
This, in turn, requires obtaining a general population sample;
one which is at least as large and which contains all the
demographics of interest.

In this case study, the survey sample is one year's worth of data
from is Cooperative Congressional Election Study (CCES),
a bi-annual survey of of 60,000 American adults from
all 50 states and the District of Columbia.
The CCES is designed to capture 
the views and experiences of Amercians with respect to Congress
and congressional (and presidential) elections.
While this survey sample is representative at the national level
it is not large enough to provide a representative sample for sub-populations
of interest.
The population sample is the 2018 American Community Survey, the
annual micro-census conducted by the US Census Bureau.  It is designed
to be a representative sample of the US population as a whole; data is
collected from 3 million households.

Big data is precious data, but a researcher's time is even more precious.
A secondary goal of this case study is showing how to work effectively
and efficiently with large samples to produce replicable results.

### Survey sample: the 2018 CCES Common Content Dataset

The 2018 CCES Common Content Dataset is 
available from the Harvard Dataverse project:
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K
The raw survey data is distributed in several from the Harvard Dataverse
as  [cces18_common_vv](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/C1YVSF&version=6.0).
Although the tabular versions of the data file are smaller, we download the
Stata file, as it contains metadata on all the questions; this is critical to
being able to align the CCES sample to the US Census population sample.
We also download the [2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
which describes the survey questions, labels, and responses.

We load the data into R using the [`readstata13`](https://sjewo.github.io/readstata13/) library function
[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) which returns a data.frame with attributes.
```{r}
library(readstata13)
cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
```

Fist we check the type, shape and contents of the sample.
```{r}
str(cces_2018, max.level=2, list.len=12, give.attr=FALSE)
dim(cces_2018)
names(cces_2018)[1:12]
```

The Common Content portion of the survey consists of questions asked of all subjects.
Profile questions provide demographic information,
and opinion questions provide outcomes of interest.
The demographic variables include
gender, education level, employment status, race, and geographic location.
```{r}
table(cces_2018$gender[,drop=TRUE])
table(cces_2018$educ[,drop=TRUE])
table(cces_2018$race[,drop=TRUE])
```
The political opinion questions on the survey are labelled "Abortion", "Immigration", "Taxes", "Health Care".
```{r}
attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]]
health_care_1 <- which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]
table(cces_2018[[health_care_1]][, drop=TRUE])
```

Of the 525 columns in the survey, we are interested in the set of political opinion questons
and a subset of the demographics.  Although not strictly necessary, we create a new dataframe
which contains just these columns of interest.
The survey metadata includes column names, and question and response labels.
We use the column names to select demographics and column labels to select political opinion questions.
```{r}
col_labels <- attr(cces_2018, "var.labels")
col_names <- attr(cces_2018, "names")
demo_cols <- c(1, which(col_names == "birthyr"),
              which(col_names == "gender" ),
              which(col_names == "educ"),
              which(col_names == "race"),
              which(col_names == "hispanic"),
              which(col_names == "marstat"),
              which(col_names == "region"),
              which(col_names == "employ"),
              which(col_names == "hadjob"),
              which(col_names == "inputstate"))
outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
                 which(startsWith(col_labels, "Immigration --")),
                 which(startsWith(col_labels, "Taxes --")),
                 which(startsWith(col_labels, "Health Care --")),
                 which(startsWith(col_labels, "President 20")))
col_labels[demo_cols]
col_labels[outcome_cols]
keep_cols <- sort(c(demo_cols, outcome_cols))
cces_2018_subset <- subset(cces_2018[keep_cols])
```

We inspect the subset to make sure that we have captured the question and response names.
```{r}
str(cces_2018_subset, max.level=2, list.len=12, give.attr=FALSE)
```

The CCES data is roughly representative of the state-level political sentiment as a whole; but if we seek to understand certain subsets of the population,
the sample may well be unbalanced.
There are only 60,000 survey responses,
but there are over 20 million possible demographic subsets of interest,
given all possible combinations of gender, education,
race, marital status, employment, birthyear, and state of residence (51).

```{r}
cat(format((length(table(cces_2018_subset$gender[,drop=TRUE])) *
length(table(cces_2018_subset$educ[,drop=TRUE])) *
length(table(cces_2018_subset$race[,drop=TRUE])) *
length(table(cces_2018_subset$marstat[,drop=TRUE])) *
length(table(cces_2018_subset$employ[,drop=TRUE])) *
length(table(cces_2018_subset$birthyr[,drop=TRUE])) *
length(table(cces_2018_subset$inputstate[,drop=TRUE]))), big.mark=",",scientific=FALSE))
```

In order to correct this imbalance, we must poststratify to the general population.

### Population sample:  the 2018 American Community Survey

The US Census data is the largest and most representative population sample we have acces to.
A full census is conducted every 10 years.
The American Community Survey (ACS) is a micro-census conducted every year by the Census Bureau
over 1% of the households in the US, i.e., it has 3 million respondants (as opposed to 300 million in the full census).

The [Integrated Public Use Microdata Series (IPUMS)](https://ipums.org/what-is-ipums) (@ipums2020) is a service run by the University of Minnesota
which allows easy access to census and survey data from around the world.
IPUMS preserves and harmonizes US census microdata, including the American Community Survey.
To use this service, you must register at https://uma.pop.umn.edu/usa/user/new.
To get this data, we work through the IPUMS interface as follows:

* From the ipums.org website, select [IPUMS USA](https://uma.pop.umn.edu/usa/user/new) and then click on “Get Data”.

* Click the "SELECT SAMPLES" button and select the 2018 ACS survey, then click on "SUBMIT SAMPLE".

* Select the variables that will be included in our poststratification table.
The survey variables are categorized by HOUSEHOLD (household-level variables) and PERSON (individual-level variables).
   + On PERSON > DEMOGRAPHIC select SEX and AGE
   + On PERSON > RACE, ETHNICITY, AND NATIVITY select RACE, HISPAN, and CITIZEN
   + On PERSON > EDUCATION select EDUC
   + On HOUSEHOLD > GEOGRAPHIC select STATEFIP

* The selected variables are added to a "data cart", along with a set of preselected variables.
At "checkout", we uncheck the preselected variables, and then click on the "CREATE DATA EXTRACT" button.
The following screenshot shows our selections, outlined
in red.

```{r ipumsExtract, echo=FALSE, out.width = "75%", fig.cap="IPUMS requested ACS variables"}
library(knitr)
include_graphics("ipums_acs_2018_match_cces.png")
```

* As even the US Census microdata files are extremely large, the IPUMS service queues this request for offline processing.
When the data has been assembled, IPUMS sends an email with a link to the download page, shown below.

```{r ipumsDownload, echo=FALSE, out.width = "75%", fig.cap="IPUMS service download page"}
library(knitr)
include_graphics("ipums_download.png")
```

You must download the data file and the DDI file for the codebook, as well as the R command file.
The DDI file is an XML file which documents your data extract.  If you click on this link,
the DDI file displays in the browser; you must download this file as XML.
The IPUMS website provides [detailed instructions](https://usa.ipums.org/usa/extract_instructions.shtml).
Both the data file and xml download file have the same base name;
if you rename them, you must rename both of them accordingly.

The following script is based on the R command script provided by IPUMS,
which requires the [`ipumsr` package](https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums.html).
This package provides methods to read, extract, inspect, and otherwise query the downloaded sample.
It also provides an excellent series of vignettes which describe how to work with the data.
We recommend starting here: http://tech.popdata.org/ipumsr/articles/ipums.html

```{r}
# NOTE: To load data, you must download both the extract's data and the DDI
# and also set the working directory to the folder with these files (or change the path below).

if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

# download dataset from IPUMS
# xml and dat file should have same name, be in same directory
acs_2018_subset_ddi <- read_ipums_ddi("usa_00007.xml")
acs_2018_subset <- read_ipums_micro(acs_2018_subset_ddi)
# save as R objects
save(acs_2018_subset, acs_2018_subset_ddi, file= "acs_2018_subset.RData")
```

#### Inspecting the ACS sample

The `ipumsr` package uses several [tidyverse](https://www.tidyverse.org/) packages
including [`dplyr`](https://dplyr.tidyverse.org/) and [`haven`](https://haven.tidyverse.org/).
The [`read_ipums_micro` function](http://tech.popdata.org/ipumsr/reference/read_ipums_micro.html)
returns a single [`tbl_df`](https://tibble.tidyverse.org/reference/tbl_df-class.html) data frame, i.e., a "tibble".
```{r}
library(ipumsr)
load("acs_2018_subset.RData")
str(acs_2018_subset, max.level=2, list.len=6)
dim(acs_2018_subset)
names(acs_2018_subset)
```
We selected variables "RACE" and "HISPAN"; the extract contains two columns for each:
"RACE" and "RACED", and "HISPAN" and "HISPAND".  The "D" columns provide an extremely fine-grained
set of categories for race and hispanic origin.  As these go far beyond the categories from
the CCES which are provided, we ignore them.

The columns of this tibble are  [labelled](https://haven.tidyverse.org/reference/labelled.html) vectors.
Labelled vectors can have text labels associated with specific values.
To see how this works, we compare the values and labels for `AGE`.
The 2018 ACS data spans ages 0 through 96,
of these, only ages 0 and 90 have labels.
```{r}
length(table(acs_2018_subset$AGE))
max(acs_2018_subset$AGE)
```

The function [`as_factor`](https://haven.tidyverse.org/reference/as_factor.html)
allows us to treat labelled vectors as factors.

### Aligning the Samples with MRP-KIT

We need to align our demographics of interest; we will do so column-by-column.

We note that information which must be aggregated from combinations of answers
given to a series of questions in order to match demographics in the ACS
is beyond the scope of what can be done with MRP-KIT, e.g.,
it doesn't provide functions for combining survey variables.

Also problematic: creating aggregate categories, e.g., binning age by decade.

#### Variable CCES "gender" / ACS "sex"

Both the CCES and ACS samples have variables for the sex of the respondant:
in the CCES this is named "gender" and in the ACS this is named "SEX".
In both, the response labels are "Male", "Female".
Creating a mapping between the two is straightforward:
we create a new SurveyQuestion object with name "sex",
then create a mappng from the CCES to the ACS,
both for column names and response labels
where for both mappings, the first element
is the survey sample name/label, and the second element
is the population sample name/label.

```{r}
library(mrpkit)
names(table(cces_2018_subset$gender))
print(ipums_val_labels(acs_2018_subset$SEX))
q_sex <- SurveyQuestion$new(name="sex",
      col_names=c("gender", "SEX"),
      values_map=list("Male" = "Male","Female" = "Female"))
print(q_sex)
```

#### Variable CCES "inputstate" / ACS "STATEFIP"

We examine the state names from both the CCES and ACS; the same labels are used in both surveys.
```{r}
# CCES data is a factor, levels are the labels
inputstate <- levels(cces_2018_subset$inputstate[,drop=TRUE])

# ACS data are structured objects, ipums_val_labels returns a tibble
STATEFIP <- ipums_val_labels(acs_2018_subset$STATEFIP)[[2]][1:51]

cat(inputstate[1:5], sep=", "); cat(STATEFIP[1:5], sep=", ")

table(inputstate == STATEFIP)
```

We need to create a SurveyQuestion object, as above,  but it would be
extremely tedious to write out a pairlist, e.g.:
```
values_map=list("Alabama" = "Alabama", ..., "Wyoming" = "Wyoming")
```
[R Pairlists](https://github.com/hadley/r-internals/blob/master/pairlists.md) are
a kind of linked list data structure where each element in the list has a name.
In order to create a pairlist programmatically, we use the R function `setNames`,
which takes as its first argument an object to be labelled
and as its second argument the corresponding character vector of labels,
i.e., the expression `as.list(c("value1", "value2"),  c("tag1", "tag2")`
is equivalent to the expression `list("tag1" = "value1", "tag2" = "value2").
```{r}
mapping = as.list(setNames(STATEFIP, inputstate))
mapping[1:3]
```

Having established the values map, we create the corresponding SurveyQuestion object.
We use accessor methods `name()`, `col_names()` and `values()` to check our work.
```{r}
q_state <- SurveyQuestion$new(name = "state",
	col_names = c("inputstate", "STATEFIP"),
	values_map= mapping)
q_state$name()
q_state$col_names()
head(q_state$values())
```