### 2021-06-30 case-study/Munging_CCES.Rmd
checkpointing case study

---
title: "Processing large surveys with package MRP-KIT"
author: "Mitzi Morris"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE, prompt=TRUE)
```

<style type="text/css">
.table { width: 85%; }
div.quote-container blockquote {  font-size: small; }
</style>

Multi-level Regresssion and Postrativication (MRP)
is a two-step process for correcting model estimates
obtained from a survey sample when the survey population is not
the same as the target population over which we wish to do inference.
The first step is the doing a multilevel regression on the survey sample;
the second step is poststratification to the target population.
Conceptually, MRP is relatively simple.
In practice, it requires more than twice as much data processing.
In addition to managing the survey data, you must also obtain
and manage the target population data, in addition you must
develop a mapping between information in these two datasets.
In order to produce verifiable, replicable results, it is
critical that this mapping be documented and scripted.

The goal of this case study is to show how the MRP-KIT package
can facilitate this process for large-scale surveys.
By large-scale surveys, we mean surveys which have both
a large number of respondants and detailed demographic profiles for each respondant.
The larger the number of demographics, the larger the number
of possible sub-populations of interest and the more difficult it becomes
to collect a representative sample for all such subpopulations,
thus even well-designed surveys can require use of MRP.

In this case study, the survey sample is one year's worth of data
from is Cooperative Congressional Election Study (CCES),
a bi-annual survey of of 60,000 American adults from
all 50 states and the District of Columbia.
The CCES is designed to capture 
the views and experiences of Amercians with respect to Congress
and congressional (and presidential) elections.
While this survey sample is representative at the national level
it is not large enough to provide a representative sample for sub-populations
of interest.
The population sample is the 2018 American Community Survey (ACS), the
annual micro-census conducted by the US Census Bureau.  It is designed
to be a representative sample of the US population as a whole; data is
collected from 3 million households.

Big data is precious data, but a researcher's time is even more precious.
A secondary goal of this case study is showing how to work effectively
and efficiently with large samples to produce replicable results.

### Survey sample: the 2018 CCES Common Content Dataset

The 2018 CCES Common Content Dataset is 
available from the Harvard Dataverse project:
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K
The raw survey data is distributed in several formats from the Harvard Dataverse
as  [cces18_common_vv](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/C1YVSF&version=6.0).
While the tabular versions of the data file are smaller, the Stata version `cces18_common_vv.dta`
contains metadata on all the questions: the question name, label, description, response labels and values.
The corresponding 
[2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
provides a human-readable version of this metadata as well.

```{r ccesDownload, echo=FALSE, out.width = "75%", fig.cap="Harvard Dataverse download page"}
library(knitr)
include_graphics("dataverse_cces_2018_download.png")
```

We download the Stata file because we will need the metatdata to map the CCES sample to the ACS using MRP-KIT,
then use the [`readstata13`](https://sjewo.github.io/readstata13/) library function
[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) to load the Stata file into R.
The result is an R data.frame with attributes containing the metadata.
```{r}
library(readstata13)
cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
```

Fist we check the type, shape and contents of the sample.
```{r}
str(cces_2018, max.level=2, list.len=12, give.attr=FALSE)
dim(cces_2018)
names(cces_2018)[1:12]
```

The Common Content portion of the survey consists of questions asked of all subjects.
Profile questions provide demographic information,
and opinion questions provide outcomes of interest.
The demographic variables include
gender, education level, race, and geographic location, among others.
```{r}
table(cces_2018$gender[,drop=TRUE])
table(cces_2018$educ[,drop=TRUE])
table(cces_2018$race[,drop=TRUE])
```
The political opinion questions on the survey are labelled "Abortion", "Immigration", "Taxes", "Health Care".
```{r}
attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]]
health_care_1 <- which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]
table(cces_2018[[health_care_1]][, drop=TRUE])
```

Of the 525 columns in the survey, we are interested in the set of political opinion questons
and a subset of the demographics.  Although not strictly necessary, we create a new dataframe
which contains just these columns of interest.
The survey metadata includes column names, and question and response labels.
We use the column names to select demographics and column labels to select political opinion questions.
```{r}
col_labels <- attr(cces_2018, "var.labels")
col_names <- attr(cces_2018, "names")
demo_cols <- c(1, which(col_names == "birthyr"),
              which(col_names == "gender" ),
              which(col_names == "educ"),
              which(col_names == "race"),
              which(col_names == "hispanic"),
              which(col_names == "marstat"),
              which(col_names == "inputstate"))
outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
                 which(startsWith(col_labels, "Immigration --")),
                 which(startsWith(col_labels, "Taxes --")),
                 which(startsWith(col_labels, "Health Care --")),
                 which(startsWith(col_labels, "President 20")))
col_labels[demo_cols]
col_labels[outcome_cols]
keep_cols <- sort(c(demo_cols, outcome_cols))
cces_2018_subset <- subset(cces_2018[keep_cols])
```

We inspect the subset to make sure that we have captured the question and response names.
```{r}
str(cces_2018_subset, max.level=2, list.len=12, give.attr=FALSE)
```

The CCES data is roughly representative of the state-level political sentiment as a whole,
but if we seek to understand certain subsets of the population,
the sample may well be unbalanced.
There are only 60,000 survey responses,
but there are over 2 million
possible combinations of gender, education,
race, marital status, birthyear, and state of residence.
If we bin ages by decade, there are a mere 200,000 combinations,
binning states by region further reduces this to 16,000, etc., etc.
However, if we wish to consider state-level subpopulations of interest,
the survey data is likely to be sparse and not representative of the general population.
To address the former we do multi-level modeling; for the latter, poststratification.

### Population sample:  the 2018 American Community Survey

The US Census data is the largest representative population sample that we have access to.
A full census is conducted every 10 years.
The American Community Survey (ACS) is a micro-census conducted every year by the Census Bureau
over 1% of the households in the US, i.e., it has 3 million respondants (as opposed to 300 million in the full census).

The [Integrated Public Use Microdata Series (IPUMS)](https://ipums.org/what-is-ipums) (@ipums2020) is a service run by the University of Minnesota
which allows easy access to census and survey data from around the world.
IPUMS preserves and harmonizes US census microdata, including the American Community Survey.
To use this service, you must register at https://uma.pop.umn.edu/usa/user/new.
To get this data, we work through the IPUMS interface as follows:

* From the ipums.org website, select [IPUMS USA](https://uma.pop.umn.edu/usa/user/new) and then click on “Get Data”.

* Click the "SELECT SAMPLES" button and select the 2018 ACS survey, then click on "SUBMIT SAMPLE".

* Select the variables that will be included in our poststratification table.
The survey variables are categorized by HOUSEHOLD (household-level variables) and PERSON (individual-level variables).
   + On PERSON > DEMOGRAPHIC select SEX and AGE
   + On PERSON > RACE, ETHNICITY, AND NATIVITY select RACE, HISPAN, and CITIZEN
   + On PERSON > EDUCATION select EDUC
   + On HOUSEHOLD > GEOGRAPHIC select STATEFIP

* The selected variables are added to a "data cart", along with a set of preselected variables.
At "checkout", we uncheck the preselected variables, and then click on the "CREATE DATA EXTRACT" button.
The following screenshot shows our selections, outlined
in red.

```{r ipumsExtract, echo=FALSE, out.width = "75%", fig.cap="IPUMS requested ACS variables"}
library(knitr)
include_graphics("ipums_acs_2018_match_cces.png")
```

The IPUMS service queues this request for offline processing.
When the data has been assembled, IPUMS sends an email with a link to the download page, shown below.

```{r ipumsDownload, echo=FALSE, out.width = "75%", fig.cap="IPUMS service download page"}
library(knitr)
include_graphics("ipums_download.png")
```

You must download the data file and the DDI file for the codebook, as well as the R command file.
The IPUMS website provides [detailed instructions](https://usa.ipums.org/usa/extract_instructions.shtml) on how to do this.
The DDI file is an XML file which documents your data extract.  If you click on this link,
the DDI file displays in the browser; you must download this file as XML.
Both the data file and xml download file have the same base name;
if you rename them, you must rename both of them accordingly.

The following script is based on the R command script provided by IPUMS,
which requires the [`ipumsr` package](https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums.html).
This package provides methods to read, extract, inspect, and otherwise query the downloaded sample.
It also provides an excellent series of vignettes which describe how to work with the data.
We recommend starting here: http://tech.popdata.org/ipumsr/articles/ipums.html

```{r}
# NOTE: To load data, you must download both the extract's data and the DDI
# and also set the working directory to the folder with these files (or change the path below).

if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

# download dataset from IPUMS
# xml and dat file should have same name, be in same directory
acs_2018_subset_ddi <- read_ipums_ddi("usa_00007.xml")
acs_2018_subset <- read_ipums_micro(acs_2018_subset_ddi)
# save as R objects
save(acs_2018_subset, acs_2018_subset_ddi, file= "acs_2018_subset.RData")
```

#### Inspecting the ACS sample

The `ipumsr` package uses several [tidyverse](https://www.tidyverse.org/) packages
including [`dplyr`](https://dplyr.tidyverse.org/) and [`haven`](https://haven.tidyverse.org/).
The [`read_ipums_micro` function](http://tech.popdata.org/ipumsr/reference/read_ipums_micro.html)
returns a single [`tbl_df`](https://tibble.tidyverse.org/reference/tbl_df-class.html) data frame, i.e., a "tibble".
```{r}
library(ipumsr)
load("acs_2018_subset.RData")
str(acs_2018_subset, max.level=2, list.len=6)
dim(acs_2018_subset)
names(acs_2018_subset)
```
We selected variables "RACE" and "HISPAN"; the extract contains two columns for each:
"RACE" and "RACED", and "HISPAN" and "HISPAND".  The "D" columns provide an extremely fine-grained
set of categories for race and hispanic origin.  As these go far beyond the categories from
the CCES which are provided, we ignore them.

The columns of this tibble are  [labelled](https://haven.tidyverse.org/reference/labelled.html) vectors.
Labelled vectors can have text labels associated with specific values.
To see how this works, we compare the values and labels for `AGE`.
The 2018 ACS data spans ages 0 through 96,
of these, only ages 0 and 90 have labels.
The function [`ipums_val_labels`](https://tech.popdata.org/ipumsr/reference/ipums_var_info.html)
allows us to inspect these labels.
```{r}
length(table(acs_2018_subset$AGE))
ipums_val_labels(acs_2018_subset$AGE)
```

The function [`as_factor`](https://haven.tidyverse.org/reference/as_factor.html)
allows us to treat labelled vectors as factors.
See the vignette https://tech.popdata.org/ipumsr/reference/ipums_var_info.html for more details.

### Aligning the Samples with MRP-KIT

We use MRP-KIT to map demographics of interest on a column-by-column basis.^[
If information is spread across multiple columns in either the survey sample
or the target population sample, it must be combined into a single column's
worth of information as MRP-KIT doesn't provide functions to do this.
*Note: all pre-processing should be documented, scripted, and **tested**!*]

In the following sections we examine each variable in turn.
This is a tedious process, therefore, we provide the following
summary of the demographic construct and the kinds of challenge each mapping presents

Construct | CCES | ACS | Challenge
--------- | ---- | --- | ---------
Sex | gender | SEX | (none)
State | inputstate | STATEFIP | specify 1-1 mapping programatically
Marital status | marstat | MARST | choice of mapping
Education | educ | EDUC | choice of mapping
Race/Hispanic ethnicity | race, hispan | RACE, HISPAN | cannot align CCES race to ACS RACE
Age | birthyr | AGE | different units, too fine-grained

#### Trivial mapping: respondant sex, 2 categories

The CCES and ACS samples record the sex of the respondant;
in the CCES this is named "gender" and in the ACS this is named "SEX".
In both, the response labels are "Male", "Female".
Creating a mapping between the two is straightforward:
we create a new SurveyQuestion object with name "sex",
then create a mappng from the CCES to the ACS,
both for column names and response labels
where for both mappings, the first element
is the survey sample name/label, and the second element
is the population sample name/label.

```{r}
library(mrpkit)
names(table(cces_2018_subset$gender))
ipums_val_labels(acs_2018_subset$SEX)
q_sex <- SurveyQuestion$new(name="sex",
      col_names=c("gender", "SEX"),
      values_map=list("Male" = "Male","Female" = "Female"))
q_sex
```

#### 1-1 mapping on a long list: state of residence, 51 categories

We examine the state names from both the CCES and ACS; the same labels are used in both surveys.
```{r}
# CCES data is a factor, levels are the labels
inputstate <- levels(cces_2018_subset$inputstate[,drop=TRUE])

# ACS data are structured objects, ipums_val_labels returns a tibble
STATEFIP <- ipums_val_labels(acs_2018_subset$STATEFIP)[[2]][1:51]

cat(inputstate[1:5], sep=", "); cat(STATEFIP[1:5], sep=", ")

table(inputstate == STATEFIP)
```

We need to create a SurveyQuestion object, as above,  but it would be
extremely tedious to write out a pairlist, e.g.:
```
values_map=list("Alabama" = "Alabama", ..., "Wyoming" = "Wyoming")
```
[R Pairlists](https://github.com/hadley/r-internals/blob/master/pairlists.md) are
a kind of linked list data structure where each element in the list has a name.
In order to create a pairlist programmatically, we use the R function `setNames`,
which takes as its first argument an object to be labelled
and as its second argument the corresponding character vector of labels,
i.e., the expression `as.list(c("value1", "value2"),  c("tag1", "tag2")`
is equivalent to the expression `list("tag1" = "value1", "tag2" = "value2").
```{r}
mapping = as.list(setNames(STATEFIP, inputstate))
mapping[1:3]
```

Having established the values map, we create the corresponding SurveyQuestion object.
We use accessor methods `name()`, `col_names()` and `values()` to check our work.
```{r}
q_state <- SurveyQuestion$new(name = "state",
	col_names = c("inputstate", "STATEFIP"),
	values_map= mapping)
q_state$name()
q_state$col_names()
head(q_state$values())
```

#### Choice of mapping: education, marital status

The CCES and ACS samples record the education level of the respondant
in variables `educ` and `EDUC`, respectively.
Both the CCES and ACS record the highest level of education completed,
but the CCES records the degree obtained whereas the ACS records the year completed.
```{r}
names(table(cces_2018_subset$educ))
ipums_val_labels(acs_2018_subset$EDUC)
```
We choose to align these categories as follows:

CCES | ACS
---- | ---
"No HS" | "Nursery school to grade 4"
"No HS" | "Grade 5, 6, 7, or 8"
"No HS" | "Grade 9"
"No HS" | "Grade 10"
"No HS" | "Grade 11"
"High School Graduate" | "Grade 12"
"Some College" | "1 year of college"
"2-year" | "2 years of college"
"2-year" | "3 years of college"
"4-year" | "4 years of college"
"Post-grad" | "5+ years of college"
"skipped" | "N/A or no schooling"
"not asked" | "N/A or no schooling"

Alternative mappings are equally viable.
With MRP-KIT, the choice of mappings is recorded and
easily accessible.
```{r}
educ_mapping = list(
    "No HS" = "Nursery school to grade 4",
    "No HS" = "Grade 5, 6, 7, or 8",
    "No HS" = "Grade 9",
    "No HS" = "Grade 10",
    "No HS" = "Grade 11",
    "High School Graduate" = "Grade 12",
    "Some College" = "1 year of college",
    "2-year" = "2 years of college",
    "2-year" = "3 years of college",
    "4-year" = "4 years of college",
    "Post-grad" = "5+ years of college",
    "skipped" = "N/A or no schooling",
    "not asked" = "N/A or no schooling"
)
q_educ <- SurveyQuestion$new(name = "educ",
    col_names = c("educ", "EDUC"),
    values_map= educ_mapping)
q_educ$name()
q_educ$col_names()
head(q_educ$values())
```

As with eduction, the CCES and ACS samples record the marital status
of the respondant differently and we have to use our judgement as to how best
to align the two.
```{r}
names(table(cces_2018_subset$marstat))
ipums_val_labels(acs_2018_subset$MARST)
```

CCES | ACS
---- | ---
"Married" | "Married, spouse present"
"Married" | "Married, spouse absent"
"Separated" | "Separated"
"Divorced" | "Divorced"
"Widowed" | "Widowed"
"Never married" | "Never married"
"Domestic / civil partnership" | "Never married"
"skipped" | "Never married"
"not asked" | "Never married"

We note that for the 2018 CCES survey, the marital status of all respondants is known; nonetheless,
we provide a complete mapping.

```{r}
marstat_mapping = list(
    "Married" = "Married, spouse present",
    "Married" = "Married, spouse absent",
    "Separated" = "Separated",
    "Divorced" = "Divorced",
    "Widowed" = "Widowed",
    "Never married" = "Never married",
    "Domestic / civil partnership" = "Never married",
    "skipped" = "Never married",
    "not asked" = "Never married"
)
q_marstat <- SurveyQuestion$new(name = "marstat",
    col_names = c("marstat", "MARST"),
    values_map= marstat_mapping)
q_marstat$name()
q_marstat$col_names()
head(q_marstat$values())
```

#### Information split across survey columns:  race and ethnicity.

We ignore the politics of race and racial categories
and instead focus on how best to map the CCES information to the ACS information.
Both surveys have multiple variables which provide information about race and ethnicity,
and both record race and hispanic origin separately.
However the CCES treats hispanic origin as race as well as ethnic origin; the ACS does not.
```{r}
names(table(cces_2018_subset$race))
names(table(cces_2018_subset$hispan))
ipums_val_labels(acs_2018_subset$RACE)
ipums_val_labels(acs_2018_subset$HISPAN)
```
In this case, munging the raw survey data is unavoidable.
For each survey we create a new variable which blends race and hispanic ethnicity.
We are scripting everyting, so we add this to our processing scripts.