### 2023-06-29 README.Rmd
use readme.Rmd to generate readme.md
this allows us to execute code in the readme for demonstration

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mrpkit

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mrpkit)](https://CRAN.R-project.org/package=mrpkit)
[![R-CMD-check](https://github.com/lauken13/mrpkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lauken13/mrpkit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/lauken13/mrpkit/branch/master/graph/badge.svg)](https://app.codecov.io/gh/lauken13/mrpkit?branch=master)
<!-- badges: end -->

**NOTE: This package is still a work in progress and is yet not released or officially supported**

For students and researchers who are comfortable with at least `glm()` and want
to conduct multilevel regression with post-stratification (MRP), `mrpkit`
provides a reproducible, opinionated, and highly structured workflow. Unlike
writing all the code yourself, using `mrpkit` proactively addresses many common
issues, and makes it possible for folks who are new to MRP to quickly conduct
their first analysis.

The package assists in setting up the survey data and relationships between
different variables in the sample and the population. From there, a substantial
amount of data cleaning is automated, saving time and reducing the risk of
coding errors. The package has native support for multilevel binomial and
Bernoulli models fit with lme4 and Stan (via brms and rstanarm) and also allows
for the use of custom modeling functions. Summaries and simple visualizations of
the resulting post-stratified estimates are provided.

## Installation

You can install the development version of `mrpkit` from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("lauken13/mrpkit")
```

### License

`mrpkit` is licensed under an MIT license. See the `LICENSE.md` file.

## Example

```{r example}
library(mrpkit)

# Some fake survey data for demonstration
head(shape_survey)

# Create SurveyData object for the sample
box_prefs <- SurveyData$new(
  data = shape_survey,
  questions = list(
    age = "Please identify your age group",
    gender = "Please select your gender",
    vote_for = "Which party did you vote for in the 2018 election?",
    y = "If today is the election day, would you vote for the Box Party?"
  ),
  responses = list(
    age = levels(shape_survey$age),
    gender = levels(shape_survey$gender),
    # Here we use a dataframe for the responses because the levels in the data are abridged versions of the actual responses
    # This can be useful when surveys have brief/non descriptive responses.
    vote_for = data.frame(data = levels(shape_survey$vote_for),
    asked = c("Box Party Faction A", "Box Party Faction B", "Circle Party Coalition", "Circle Party")),
    y = c("no","yes")
  ),
  weights = "wt",
  design = list(ids =~1)
)
box_prefs$print()
box_prefs$n_questions()

# Some fake population data for demonstration
head(approx_voters_popn)

# Create SurveyData object for the population
popn_obj <- SurveyData$new(
  data = approx_voters_popn,
  questions = list(
    age_group = "Which age group are you?",
    gender = "Gender?",
    vote_pref = "Which party do you prefer to vote for?"
  ),
  # order doesn't matter (gender before age here) because
  # the list has the names of the variables
  responses = list(
    gender = levels(approx_voters_popn$gender),
    age_group = levels(approx_voters_popn$age_group),
    vote_pref = levels(approx_voters_popn$vote_pref)
  ),
  weights = "wt"
)
popn_obj$print()

# Create the QuestionMap objects mapping each question between the
# survey and population dataset
q_age <- QuestionMap$new(
  name = "age",
  col_names = c("age","age_group"),
  values_map = list(
    "18-25" = "18-35", "26-35" = "18-35","36-45" = "36-55",
    "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
  )
)
print(q_age)

q_party_pref <- QuestionMap$new(
  name = "party_pref",
  col_names = c("vote_for","vote_pref"),
  values_map = list("Box Party" = "BP",  "BP" = "BP","Circle Party" = "CP", "CP" = "CP")
)
q_gender <- QuestionMap$new(
  name = "gender",
  col_names = c("gender", "gender"),
  values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
)

# Create SurveyMap object adding all questions at once
ex_map <- SurveyMap$new(
  sample = box_prefs,
  population = popn_obj,
  q_age,
  q_party_pref,
  q_gender
)
print(ex_map) # or ex_map$print()

# Or can add questions incrementally
ex_map <- SurveyMap$new(sample = box_prefs, population = popn_obj)
print(ex_map)

ex_map$add(q_age, q_party_pref)
print(ex_map)

ex_map$add(q_gender)
print(ex_map)

# Create the mapping between sample and population
ex_map$mapping()

# Create the poststratification data frame using all variables in the mapping
# (alternatively, can specify particular variables, e.g. tabulate("age"))
ex_map$tabulate()

# Take a peak at the poststrat data frame
head(ex_map$poststrat_data())

# Fit regression model using rstanarm (returns a SurveyFit object)
fit_1 <- ex_map$fit(
  fun = rstanarm::stan_glmer,
  formula = y ~ (1|age) + (1|gender),
  family = "binomial",
  seed = 1111,
  chains = 1, # just to keep the example fast and small
  refresh = 0 # suppress printed sampling iteration updates
)

# To use lme4 or brms instead of rstanarm you would use: 
# Example lme4 usage
# fit_2 <- ex_map$fit(
#   fun = lme4::glmer,
#   formula = y ~ (1|age) + (1|gender),
#   family = "binomial"
# )

# Example brms usage
# fit_3 <- ex_map$fit(
#   fun = brms::brm,
#   formula = y ~ (1|age) + (1|gender),
#   family = "bernoulli",
#   seed = 1111
# )

# Predicted probabilities
# returns matrix with rows for poststrat cells, cols for posterior draws
poststrat_estimates <- fit_1$population_predict()

# Compute and summarize estimates by age level and party preference
estimates_by_age <- fit_1$aggregate(poststrat_estimates, by = "age")
estimates_by_party <- fit_1$aggregate(poststrat_estimates, by = "party_pref")

fit_1$summary(estimates_by_age)
fit_1$summary(estimates_by_party)

# Plot estimates
fit_1$plot(estimates_by_party)

fit_1$plot(estimates_by_age)

fit_1$plot(estimates_by_age, additional_stats = "none")
fit_1$plot(estimates_by_age, additional_stats = "wtd")
fit_1$plot(estimates_by_age, additional_stats = "raw")
fit_1$plot(estimates_by_age, additional_stats = c("wtd","raw","mrp"))

# Compute and summarize the population estimate
estimates_popn <- fit_1$aggregate(poststrat_estimates)
fit_1$summary(estimates_popn)

# Plot population estimate
fit_1$plot(estimates_popn)
fit_1$plot(estimates_popn, additional_stats = "none")
fit_1$plot(estimates_popn, additional_stats = "wtd")
fit_1$plot(estimates_popn, additional_stats = "raw")
fit_1$plot(estimates_popn, additional_stats = c("wtd","raw","mrp"))
```